Geometric variational problems

### 2020-11-16 NSF2020/Description/preliminarywork.tex
yny Nov15c
RJR

\vspace{-0.1cm}
\begin{wraptable}[8]{r}{0.53\textwidth}
\vspace{-0.1cm}
\centering{
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & experiment          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
}
  \vspace{-8pt}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
\section{Preliminary work}
\label{sec:preliminary_work}
PIs RR and YNY started work on HAP modeling with Szu-Pei Fu (then PI YNY's student and now PI RR's postdoc) and collaborators in 2017. 
%
%
%The HAP project was initiated in 2017 by PI RR and YNY. The lead author
%on our first paper~\cite{Fu2018_SIAM} was Szu-Pei Fu, who was PI YNY's
%PhD student and is currently a postdoc with PI RR. The collaboration
%involved theoretical work and simulation. The theoretical work included
%model development, proving the first variation formula~\eqref{stress}
%and force-free law, and an energy-based uniqueness theorem for the
%exterior domain problem~\eqref{SL}. 
%Coauthors Andreas Kl\"ockner
%(Computer Science, UIUC, NSF CAREER \#1654756) and his PhD student Matt
%Wala provided their Quadrature by Expansion, or QBX algorithm and Fu ran
%simulations on NJIT clusters.
%
%\begin{wraptable}[11]{l}{0.43\textwidth}
%\centering{
%  \begin{tabular}{|l|l|l|l|}
%    $n$   & (rel.~err.)$_F$
%    &  (rel.~err.)$_G$ \\
%    \hline
%% 32   &   4.07466     &   1.56437\\
%% 64   &   0.24664     &   0.03476\\
%% 128  &   0.00063     &   0.00006\\
%  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
%  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
%  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
%  \end{tabular}}
%\vspace{-5pt}
%\caption{\label{tab:spectral_force} 
%\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
%  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
%  and (rel.~err.)$_G = \max_{i}
%  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
%  for force and torque respectively as a function of number of grid
%  points $n$ per particle.  The data is for $N = 5$ particles; two of
%  the particles are nearly touching at a distance 1\% of particle
%  diameter.} 
%\end{wraptable}
The results are summarized in \cite{Fu2018_SIAM}, where
small particle-number simulations successfully demonstrated gradient-driven self-assembly of
amphiphilic particles into two-dimensional micelles and bilayer membranes with realistic values of
the model parameters:  decay length $\rho=2.5$ nm (based on measurements of
hydrophobic attraction between surfactant-coated
membranes~\cite{Eriksson1989, Lin2005, Parsegian, Israelachvili80}),
particle diameter 2.0 nm (the physical monolayer thickness, 2.0 nm~\cite{Boal}),
and interfacial tension $\gamma = 4$ pN nm$^{-1}$ based on
measurements for single-component bilayer lipid
membranes~\cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson2016}.

@@

\begin{wraptable}[11]{l}{0.43\textwidth}
\centering{
  \begin{tabular}{|l|l|l|l|}
    $n$   & (rel.~err.)$_F$
    &  (rel.~err.)$_G$ \\
    \hline
% 32   &   4.07466     &   1.56437\\
% 64   &   0.24664     &   0.03476\\
% 128  &   0.00063     &   0.00006\\
  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
  \end{tabular}}
\vspace{-5pt}
\caption{\label{tab:spectral_force} 
\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
  and (rel.~err.)$_G = \max_{i}
  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
  for force and torque respectively as a function of number of grid
  points $n$ per particle.  The data is for $N = 5$ particles; two of
  the particles are nearly touching at a distance 1\% of particle
  diameter.} 
\end{wraptable}
PIs BQ, RR and YNY initiated a collaboration in Summer 2020 to
develop the hydrodynamic model of \S\ref{subsec:specific_aim_3}.  To
date, we have written solvers for the two-dimensional boundary integral
formulations of~\eqref{SL} and~\eqref{eq:stokes} that achieve
third-order accuracy up to the boundary. The specialized reciprocal
identity~\eqref{eq:reciprocal} gives spectral accuracy for force and
torque (Table~\ref{tab:spectral_force}).

\begin{wrapfigure}[]{r}{0.5\textwidth}
  \centering{\includegraphics[width=0.5\textwidth]{Figures/TankTreading.pdf}}
  \vspace{-25pt}
  \caption{\label{fig:tank_treading} \footnotesize
    Panel A: A two-dimensional
    particle-based vesicle in a shear flow (arrows); $\mu  = 1$ mPa s,
    $\dot{\gamma} = 0.5$ $\mu$s$^{-1}$. The color coding is for HAP. Panel B: The
    mean angular velocities for the outer leaflet and the inner
    respectively.  Panel C: The trajectory of the particle
    centers forms two ellipses with well-defined inclination angle
    $\Theta$. Panel D: The evolution of the inclination angle.
    Panels B and D share the same time axis.}
\end{wrapfigure}

We have simulated two-dimensional vesicles in shear flow
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate
$\dot{\gamma}$,
%(The symbol shear rate is unrelated to interfacial tension.)
and the particle-based system mimics the behavior observed in continuous
vesicles~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. The basic
pattern is that the vesicle bilayer approaches a steady tank-treading
ellipse (Figure~\ref{fig:tank_treading}C). We analyzed the simulation
data in the context of theoretical work on
tank-treading~\cite{Finken2008, PhysRevLett.106.158103}.
Jeffrey's equation predicts a
tank-treading frequency $-\dot{\gamma}/2$ (dashed line in
Figure~\ref{fig:tank_treading}B). In the particle simulation, the
tank-treading frequency corresponds to the mean of the particle angular
velocities, and these approach a value within 7\% of the theoretical
prediction (dotted and solid curves in Figure~\ref{fig:tank_treading}B).
In the low viscosity-contrast regime of our simulation, Jeffrey's
equation predicts a decrease in the ellipse's inclination angle $\Theta$
and our particle-based simulation follows this prediction
(Figure~\ref{fig:tank_treading}D), albeit with some higher frequency oscillation
coming from particle reorganization.

Analyzing for any underlying constitutive laws, the particles reorganize
so that there is an initial increase and then plateauing of vesicle
circumference. As is the case with real lipid bilayer membranes, the
particle collection permits a small amount of stretching. The
theoretical calculations give an increase of length by 3.6\% and this amount
agrees with our measured increase in length. The simulations also
suggest that the particle system behaves as a semipermeable
vesicle~\cite{323e9a2f0c58487ea82518d7a1f96485, YAO2017728}.

\begin{wrapfigure}[11]{l}{0.22\textwidth}
  \centering{\includegraphics[width=0.22\textwidth]{Figures/PW_fig5.pdf}}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction to the point that
the vesicle ruptures. The flow carries away segmented membrane patches
(Figure \ref{fig:flow_rupture}).

In conclusion, the HAP model accurately mimics the behavior of continuous
membranes, while at the same time capturing the
reconnection during the topological change of 
lipid molecules on the scales of membrane thickness,
overcoming one of the long-standing challenges of continuum modeling.

>>>>

\vspace{-0.1cm}
\begin{wraptable}[8]{r}{0.53\textwidth}
\vspace{-0.1cm}
\centering{
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & experiment          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
}
  \vspace{-8pt}
\caption{\label{tab:moduli} \footnotesize Comparison of values of elastic
  moduli from the experimental literature and derived by HAP
  simulation.} 
\end{wraptable}
\section{Preliminary work}
\label{sec:preliminary_work}
PIs RR and YNY started work on HAP modeling with Szu-Pei Fu (then PI YNY's student and now PI RR's postdoc) and collaborators in 2017. 
%
%
%The HAP project was initiated in 2017 by PI RR and YNY. The lead author
%on our first paper~\cite{Fu2018_SIAM} was Szu-Pei Fu, who was PI YNY's
%PhD student and is currently a postdoc with PI RR. The collaboration
%involved theoretical work and simulation. The theoretical work included
%model development, proving the first variation formula~\eqref{stress}
%and force-free law, and an energy-based uniqueness theorem for the
%exterior domain problem~\eqref{SL}. 
%Coauthors Andreas Kl\"ockner
%(Computer Science, UIUC, NSF CAREER \#1654756) and his PhD student Matt
%Wala provided their Quadrature by Expansion, or QBX algorithm and Fu ran
%simulations on NJIT clusters.
%
%\begin{wraptable}[11]{l}{0.43\textwidth}
%\centering{
%  \begin{tabular}{|l|l|l|l|}
%    $n$   & (rel.~err.)$_F$
%    &  (rel.~err.)$_G$ \\
%    \hline
%% 32   &   4.07466     &   1.56437\\
%% 64   &   0.24664     &   0.03476\\
%% 128  &   0.00063     &   0.00006\\
%  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
%  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
%  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
%  \end{tabular}}
%\vspace{-5pt}
%\caption{\label{tab:spectral_force} 
%\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
%  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
%  and (rel.~err.)$_G = \max_{i}
%  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
%  for force and torque respectively as a function of number of grid
%  points $n$ per particle.  The data is for $N = 5$ particles; two of
%  the particles are nearly touching at a distance 1\% of particle
%  diameter.} 
%\end{wraptable}
The results are summarized in \cite{Fu2018_SIAM}, where
small particle-number simulations successfully demonstrated gradient-driven self-assembly of
amphiphilic particles into two-dimensional micelles and bilayer membranes with realistic values of
the model parameters: decay length $\rho=2.5$ nm (based on measurements of
hydrophobic attraction between surfactant-coated
membranes~\cite{Eriksson1989, Lin2005, Parsegian, Israelachvili80}),
particle diameter $\delta = 2.0$ nm (the physical monolayer thickness~\cite{Boal}),
and interfacial tension $\gamma = 4$ pN nm$^{-1}$ (based on
measurements for single-component bilayer lipid
membranes~\cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson2016}).

@@

\begin{wraptable}[11]{l}{0.43\textwidth}
\centering{
  \begin{tabular}{|l|l|l|l|}
    $n$   & (rel.~err.)$_F$
    &  (rel.~err.)$_G$ \\
    \hline
% 32   &   4.07466     &   1.56437\\
% 64   &   0.24664     &   0.03476\\
% 128  &   0.00063     &   0.00006\\
  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
  \end{tabular}}
\vspace{-5pt}
\caption{\label{tab:spectral_force} 
\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
  and (rel.~err.)$_G = \max_{i}
  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
  for force and torque, respectively, as a function of number of grid
  points $n$ per particle. The data is for $N = 5$ particles; two of
  the particles are nearly touching at a distance 1\% of particle
  diameter.} 
\end{wraptable}
PIs BQ, RR, and YNY initiated a collaboration in Summer 2020 to
develop the hydrodynamic model of \S\ref{subsec:specific_aim_3}.  To
date, we have written solvers for the two-dimensional boundary integral
formulations of~\eqref{SL} and~\eqref{eq:stokes} that achieve
third-order accuracy up to the boundary. The specialized reciprocal
identity~\eqref{eq:reciprocal} gives spectral accuracy for force and
torque (Table~\ref{tab:spectral_force}).

\begin{wrapfigure}[]{r}{0.5\textwidth}
  \centering{\includegraphics[width=0.5\textwidth]{Figures/TankTreading.pdf}}
  \vspace{-25pt}
  \caption{\label{fig:tank_treading} \footnotesize
    Panel A: A two-dimensional,
    particle-based vesicle in a shear flow (arrows); $\mu  = 1$ mPa s and
    $\dot{\gamma} = 0.5$ $\mu$s$^{-1}$. The color coding is for HAP. Panel B: The
    mean angular velocities for the outer and the inner leaflets, 
    respectively. Panel C: The trajectory of the particle
    centers forms two ellipses with well-defined inclination angle
    $\Theta$. Panel D: The evolution of the inclination angle.
    Panels B and D share the same time axis.}
\end{wrapfigure}

We have simulated two-dimensional vesicles in shear flow
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate
$\dot{\gamma}$,
%(The symbol shear rate is unrelated to interfacial tension.)
and the particle-based system mimics the behavior observed in continuous
vesicles~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. The basic
pattern is that of a vesicle bilayer approaching a steady, tank-treading
ellipse (Figure~\ref{fig:tank_treading}C). We analyzed the simulation
data in the context of theoretical work on
tank-treading~\cite{Finken2008, PhysRevLett.106.158103}.
Jeffrey's equation predicts a tank-treading frequency $-\dot{\gamma}/2$ (dashed line in
Figure~\ref{fig:tank_treading}B). In the particle simulation, we measured
a tank-treading frequency as the mean, particle angular
velocities. The measured steady-state frequencies were within 7\% of the predicted 
frequency (dotted and solid curves in Figure~\ref{fig:tank_treading}B).
The inclincation angle of the particle-based simulation also
follows the inclination angle $\Theta$ predicted by Jeffrey's
equation (Figure~\ref{fig:tank_treading}D), albeit with some higher frequency oscillation
coming from particle reorganization.

Analyzing for any underlying constitutive laws, the particles reorganize
so that there is an initial increase and then plateau in vesicle
circumference. As is the case with real lipid bilayer membranes, the
particle collection permits a small amount of stretching. The
theoretical calculations give an increase of length by 3.6\%, and this amount
agrees with our measured increase in length. The simulations also
suggest that the particle system behaves as a semipermeable
vesicle~\cite{323e9a2f0c58487ea82518d7a1f96485, YAO2017728}.

\begin{wrapfigure}[11]{l}{0.22\textwidth}
  \centering{\includegraphics[width=0.22\textwidth]{Figures/PW_fig5.pdf}}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction to the point that
the vesicle ruptures, and the flow carries away segmented membrane patches
(Figure \ref{fig:flow_rupture}).

In conclusion, the HAP model accurately mimics the behavior of continuous
membranes while at the same time capturing the
reconnection during the topological change of 
lipid molecules on the scales of membrane thickness,
overcoming one of the long-standing challenges of continuum modeling.
